---
title: "Day 1 Exercise 4"
subtitle: "Tidying data with `tidyr`"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Tidying data with tidyr
---

```{css, echo = FALSE}
.mytask {
  background: #82CAFA;
  border: 1.2px dashed #151B54;
  color: #151B54;
  padding: 10px;
  margin: auto;
  text-align: center;
}
.mytask:before {
  content: "Task";
  font-weight: bold;
}
```

```{r setup, include=FALSE, error = TRUE}
library(learnr)
library(tidyverse)
library(dwexercise)
if("gradethis" %in% installed.packages()) {
  gradethis::gradethis_setup()
} 
tutorial_options(exercise.timelimit = 60,
                 exercise.eval = FALSE)
knitr::opts_chunk$set(error = TRUE)
```

## Welcome

* You'll be working through exercises that help you to transform the data to a tidy data form.
* Note that answers can be achieved in multiple ways. 
* You'll be working through the `frog_signal` and `tidy_abstracts` datasets once again.
* The datasets and the `tidyverse` package have been loaded for all exercises.

```{r exercise0A, exercise = TRUE}
frog_signal
```


```{r exercise0B, exercise = TRUE}
tidy_abstracts
```


## (A) Pivotting data

<div class="mytask">
1. Remove variables that were drived from other variables, i.e. `ChoiceBinary`, `StandardAverage`, `AlternativeAverage`, `PhonotaxisScore`, `ScorePreference` and `Mismatch` from the `frog_signal` data.
2. Rename `AlternativeCD`, `AlternativeCR`, `AlternativeDF`, `AlternativeRA`, and `AlternativePR` so that it does not have the prefix "Alternative".
3. Pivot data so that there are (i) a column called `call` which stores whether the call was `Standard`, `Alternative` or `Both` and (ii) a column called `response` which stores the latent response. Note: the headers `Standard1`, `Standard2`, `Standard3` contain the latent response from Standard calls; `Alternative1`, `Alternative2`, `Alternative3` contain the latent response from Alternative calls; and `TwoChoiceLatency` contains the latent response from Both. 
4. Modify `CD`, `CR`, `DF`, `RA`, and `PR` (the independent values for the Alternative call) so that the values are `NA` if the call was Standard.
5. Modify `TwoChoice` so that it has the value "Standard" when `call` was Standard, "Alternative" when `call` was Alternative, otherwise store its own value.
</div>

```{r exercise1, exercise = TRUE}
frog_signal %>% 
  select(...) %>% 
  rename_with(...) %>% 
  ...(..., 
               names_to = "call",
               values_to = "response", 
               names_transform = list(call = function(x) {
                  ... })) %>% 
  mutate(across(CD:PR, function(x) ifelse(call=="Standard", NA, x)),
         TwoChoice = ifelse(call=="Both", TwoChoice, call))
```



```{r exercise1-check}
grade_this({
  df <- frog_signal %>% 
    select(-c(ChoiceBinary:Mismatch)) %>% 
    rename_with(function(x) str_replace(x, "Alternative", ""),
                .cols = AlternativeCD:AlternativePR) %>% 
    pivot_longer(Standard1:TwoChoiceLatency, 
                 names_to = "call",
                 values_to = "response", 
                 names_transform = list(call = function(x) {
                    ifelse(x=="TwoChoiceLatency", 
                           "Both", str_replace(x, "[1-3]", ""))})) %>% 
    mutate(across(CD:PR, function(x) ifelse(call=="Standard", NA, x)),
           TwoChoice = ifelse(call=="Both", TwoChoice, call))
  pass_if_equal(df)
})
```

```{r exercise1-solution}
frog_signal %>% 
  select(-c(ChoiceBinary:Mismatch)) %>% 
  rename_with(function(x) str_replace(x, "Alternative", ""),
              .cols = AlternativeCD:AlternativePR) %>% 
  pivot_longer(Standard1:TwoChoiceLatency, 
               names_to = "call",
               values_to = "response", 
               names_transform = list(call = function(x) {
                  ifelse(x=="TwoChoiceLatency", 
                         "Both", str_replace(x, "[1-3]", ""))})) %>% 
  mutate(across(CD:PR, function(x) ifelse(call=="Standard", NA, x)),
         TwoChoice = ifelse(call=="Both", TwoChoice, call))
```

* This is quite difficult! **See below for the expected output.** 

```{r exercise1-print, echo = FALSE}
frog_signal %>% 
  select(-c(ChoiceBinary:Mismatch)) %>% 
  rename_with(function(x) str_replace(x, "Alternative", ""),
              .cols = AlternativeCD:AlternativePR) %>% 
  pivot_longer(Standard1:TwoChoiceLatency, 
               names_to = "call",
               values_to = "response", 
               names_transform = list(call = function(x) {
                  ifelse(x=="TwoChoiceLatency", 
                         "Both", str_replace(x, "[1-3]", ""))})) %>% 
  mutate(across(CD:PR, function(x) ifelse(call=="Standard", NA, x)), 
         TwoChoice = ifelse(call=="Both", TwoChoice, call))
```

<div class="mytask">
The `SpeakerPosition` was not recorded for certain experiments. Remove observations where the `SpeakerPosition` was not recorded. And then, calculate the mean of `StandardAverage` and `AlternativeAverage` for each level of `SpeakerPosition`. Transpose this summary data by having the speaker position as the header and the rows by the call type. 
</div>

```{r exercise4, exercise = TRUE}
frog_signal %>% 
  ...(!is.na(SpeakerPosition)) %>% 
  ...(SpeakerPosition) %>% 
  ...(Standard = mean(StandardAverage),
            Alternative = mean(AlternativeAverage)) %>% 
  ...(c(Standard, Alternative),
      values_to = "response",
      names_to = "call") %>% 
  ...(... = SpeakerPosition,
      ... = response)
```



```{r exercise4-check}
grade_this({
  df <- frog_signal %>% 
    filter(!is.na(SpeakerPosition)) %>% 
    group_by(SpeakerPosition) %>% 
    summarise(Standard = mean(StandardAverage),
              Alternative = mean(AlternativeAverage)) %>% 
    pivot_longer(c(Standard, Alternative),
                 values_to = "response",
                 names_to = "call") %>% 
    pivot_wider(names_from = SpeakerPosition,
                values_from = response)
  pass_if_equal(df)
})
```

```{r exercise4-solution}
frog_signal %>% 
  filter(!is.na(SpeakerPosition)) %>% 
  group_by(SpeakerPosition) %>% 
  summarise(Standard = mean(StandardAverage),
            Alternative = mean(AlternativeAverage)) %>% 
  pivot_longer(c(Standard, Alternative),
               values_to = "response",
               names_to = "call") %>% 
  pivot_wider(names_from = SpeakerPosition,
              values_from = response)
```

* Below is the expected output

```{r exercise4-print, echo = FALSE, message = FALSE}
frog_signal %>% 
  filter(!is.na(SpeakerPosition)) %>% 
  group_by(SpeakerPosition) %>% 
  summarise(Standard = mean(StandardAverage),
            Alternative = mean(AlternativeAverage)) %>% 
  pivot_longer(c(Standard, Alternative),
               values_to = "response",
               names_to = "call") %>% 
  pivot_wider(names_from = SpeakerPosition,
              values_from = response)
```


## (B) Transforming data to a tidy form

<div class="mytask">
Transform the `tidy_abstracts` data so that each cell in `authors` contain only one author by separating them out to their own row and rename the column to `author`. Add a column called `order`, located after `author`, that lists the author in order listed for the corresponding paper.
</div>

```{r exercise2, exercise = TRUE}
tidy_abstracts %>% 
  ... %>% 
  rename(author = authors) %>% 
  group_by(title) %>% 
  mutate(order = 1:n()) %>% # what do you think n() does here?
  ...(order, .after = author)
```



```{r exercise2-check}
grade_this({
  df <- tidy_abstracts %>% 
    separate_rows(authors, sep = " and ") %>% 
    rename(author = authors) %>% 
    group_by(title) %>% 
    mutate(order = 1:n()) %>% 
    relocate(order, .after = author)
  pass_if_equal(df)
})
```

```{r exercise2-solution}
tidy_abstracts %>% 
  separate_rows(authors, sep = " and ") %>% 
  rename(author = authors) %>% 
  group_by(title) %>% 
  mutate(order = 1:n()) %>% 
  relocate(order, .after = author)
```

* Below is the expected output

```{r exercise2-print, echo = FALSE, message = FALSE}
tidy_abstracts %>% 
  separate_rows(authors, sep = " and ") %>% 
  rename(author = authors) %>% 
  group_by(title) %>% 
  mutate(order = 1:n()) %>% 
  relocate(order, .after = author)
```

<div class="mytask">
`FrogID` in `frog_signal` is a 5-digit code where the first two digits indicate the year in which that female was collected and tested, while the next three digits are a unique identifier. Seperate `FrogID` to columns `year` and `id`. Modfiy the year so say "13" is the integer 2013. Output the data frame with `FrogID`, `year` and `id`.
</div>

```{r exercise3, exercise = TRUE}
frog_signal %>% 
  ... %>% 
  mutate(year = as.integer(str_c("20", year))) %>% 
  select(FrogID, year, id)
```



```{r exercise3-check}
grade_this({
  df <- frog_signal %>% 
    extract(FrogID, 
            into = c("year", "id"),
            regex = "([0-9]{2})([0-9]{3})",
            remove = FALSE) %>% 
    mutate(year = as.integer(str_c("20", year))) %>% 
    select(FrogID, year, id)
  pass_if_equal(df)
})
```

```{r exercise3-solution}
frog_signal %>% 
  extract(FrogID, 
          into = c("year", "id"),
          regex = "([0-9]{2})([0-9]{3})",
          remove = FALSE) %>% 
  mutate(year = as.integer(str_c("20", year))) %>% 
  select(FrogID, year, id)
```

* Below is the expected output

```{r exercise3-print, echo = FALSE}
frog_signal %>% 
  extract(FrogID, 
          into = c("year", "id"),
          regex = "([0-9]{2})([0-9]{3})",
          remove = FALSE) %>% 
  mutate(year = as.integer(str_c("20", year))) %>% 
  select(FrogID, year, id)
```

