---
title: "Day 1 Exercise 3"
subtitle: "Manipulating strings with `stringr`"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Manipulating strings with stringr
---

```{css, echo = FALSE}
.mytask {
  background: #82CAFA;
  border: 1.2px dashed #151B54;
  color: #151B54;
  padding: 10px;
  margin: auto;
  text-align: center;
}
.mytask:before {
  content: "Task";
  font-weight: bold;
}
```

```{r setup, include=FALSE, error = TRUE}
library(learnr)
library(tidyverse)
data(awl, package = "dwexercise")
if("gradethis" %in% installed.packages()) {
  gradethis::gradethis_setup()
} 
tutorial_options(exercise.timelimit = 60,
                 exercise.eval = FALSE)
knitr::opts_chunk$set(error = TRUE)
```

## Welcome

* This exercise will get you to work through the various functions in `stringr` and also let you practice using regular expressions. 
* You will also be using some of the `dplyr` verbs you've just learnt.
* As usual, the dataset and the `tidyverse` package (containing `dplyr` and `stringr`) have been already loaded for you.
* Remember that there are multiple ways to achieve the answer.
* You'll working will the Academic Word List (AWL). 
* The AWL was developed by Averil Coxhead as her MA thesis at the School of Linguistics and Applied Language Studies at Victoria University of Wellington, New Zealand. 
* The list contains 570 word families which were selected according to principles. 
* The list does not include words that are in the most frequent 2000 words of English.
* There list contains 10 sublist families ordered by their frequency such that sublist 1 contains the most common words in AWL with sublist 10 with the least common words in AWL. 
* You can see the data set below.

```{r exercise0, exercise = TRUE}
data(awl, package = "dwexercise")
awl
```


## (A) Word search


<div class="mytask">
How many words are there for each sublist?
</div>

```{r exercise1, exercise = TRUE}
awl %>% 
  ...(sublist) %>% 
  tally() # this is from dplyr
```

```{r exercise1-check}
grade_this({
  df <- awl %>% 
    group_by(sublist) %>% 
    tally()
  pass_if_equal(df)
})
```

```{r exercise1-solution}
awl %>% 
  group_by(sublist) %>% 
  tally()
```

<div class="mytask">
How many words are there in AWL that ends with "tion" or "tions".
</div>

```{r exercise2, exercise = TRUE}
sum(str_...(awl$word, ...))
```

```{r exercise2-check}
grade_this({
  pass_if_equal(sum(str_ends(awl$word, "tions?")))
})
```

```{r exercise2-solution}
sum(str_ends(awl$word, "tions?"))
```

<div class="mytask">
How many words are there in AWL that ends with "tion" or "tions".
</div>

```{r exercise2, exercise = TRUE}
sum(str_...(awl$word, ...))
```

```{r exercise2-check}
grade_this({
  pass_if_equal(sum(str_ends(awl$word, "tions?")))
})
```

```{r exercise2-solution}
sum(str_ends(awl$word, "tions?"))
```

<div class="mytask">
What are the shortest and the longest words in AWL that end with "ed"? Have your last output with these words in a vector ordered alphabetically. 
</div>

```{r exercise2, exercise = TRUE}
awl %>% 
  ...(...(word, "ed")) %>% 
  ...(length = ...(word)) %>% 
  ...(length %in% c(max(length), min(length))) %>% 
  arrange(word) %>% 
  pull(word)
```

```{r exercise2-check}
grade_this({
  out <- awl %>% 
    filter(str_ends(word, "ed")) %>% 
    mutate(length = str_length(word)) %>% 
    filter(length %in% c(max(length), min(length))) %>% 
    arrange(word) %>% 
    pull(word)
  
  pass_if_equal(out)
})
```

```{r exercise2-solution}
awl %>% 
  filter(str_ends(word, "ed")) %>% 
  mutate(length = str_length(word)) %>% 
  filter(length %in% c(max(length), min(length))) %>% 
  arrange(word) %>% 
  pull(word)
```


<div class="mytask">

</div>

```{r exercise3, exercise = TRUE}

```

```{r exercise3-check}
grade_this({

  
  pass_if_equal(out)
})
```

```{r exercise3-solution}

```