---
title: "Day 1 Exercise 1"
subtitle: "Knowing your `base` functions"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Revision of Base R functions.
---

```{css, echo = FALSE}
.mytask {
  background: #82CAFA;
  border: 1.2px dashed #151B54;
  color: #151B54;
  padding: 10px;
  margin: auto;
  text-align: center;
}
.mytask:before {
  content: "Task";
  font-weight: bold;
}
```


```{r setup, include=FALSE, error = TRUE}
library(learnr)
if("gradethis" %in% installed.packages()) {
  gradethis::gradethis_setup()
} 
tutorial_options(exercise.timelimit = 60,
                 exercise.eval = FALSE)
knitr::opts_chunk$set(error = TRUE)
```

## Welcome

* This exercise is a revision of your knowledge in Base R. 
* **You should complete the task _without_ loading any extra packages.**
* **Note that answers can be achieved in multiple ways.**
* The solution just shows one way to achieve the answer.
* You'll working through the `mtcars` dataset.
* Try pushing the `Run code` button. You should see the data frame output.

```{r exercise0A, exercise = TRUE}
mtcars
```

<div class="mytask">
Extract the `mpg` vector from the `mtcars` dataset.
</div>

* Note: the code has already been written for you. 

```{r exercise0B, exercise = TRUE}
mtcars[, "mpg"]
```

```{r exercise0B-check}
grade_this({
  pass_if_equal(mtcars[, "mpg"])
})
```

* Push `Submit Answer` button above and it will tell you whether you got your answer correct.
* Looks like you have gotten the hang of using this `learnr` shiny app to do the questions! Click on Next Topic to move on.

## (A) Subset with logical vectors

<div class="mytask">
Extract every odd row from `mtcars`.
</div>

```{r exercise1, exercise = TRUE}
mtcars[..., ]
```

```{r exercise1-check}
grade_this({
  pass_if_equal(mtcars[c(TRUE, FALSE), ])
})
```

```{r exercise1-solution}
mtcars[c(TRUE, FALSE), ]
```


<details><summary>Hint</summary>
You can pass a logical vector to a `data.frame` object to subset rows or columns. What happens if the vector is shorter than the number of rows (or columns)?
</details>

## (B) Deleting a column or row

<div class="mytask">
Delete the `wt` column from `mtcars`. Remember to output the data frame after you delete it. 
</div>

```{r exercise2, exercise = TRUE}
mtcars$wt <- ...
mtcars
```

```{r exercise2-check}
grade_this({
  df <- mtcars
  df$wt <- NULL
  pass_if_equal(df)
})
```

```{r exercise2-solution}
mtcars$wt <- NULL
mtcars
```

<div class="mytask">
Delete the first five rows from `mtcars`. Remember to output the data frame after you delete it. 
</div>

```{r exercise3, exercise = TRUE}
mtcars
```

```{r exercise3-check}
grade_this(
  pass_if_equal(mtcars[-c(1:5), ])
)
```

```{r exercise3-solution}
mtcars[-c(1:5), ]
```

## (C) Finding the dimension of data

<div class="mytask">
Find the number of rows for `mtcars`.
</div>

```{r exercise4, exercise = TRUE}
...(mtcars)
```

```{r exercise4-check}
grade_this(
  pass_if_equal(nrow(mtcars))
)
```

```{r exercise4-solution}
nrow(mtcars)
```

<div class="mytask">
Find the number of columns for `mtcars`.
</div>

```{r exercise5, exercise = TRUE}
...(mtcars)
```

```{r exercise5-check}
grade_this(
  pass_if_equal(ncol(mtcars))
)
```

```{r exercise5-solution}
ncol(mtcars)
```

## (D) Row name of the data

<div class="mytask">
Get the row names of the data.
</div>

```{r exercise6, exercise = TRUE}
...(mtcars)
```

```{r exercise6-check}
grade_this(
  pass_if_equal(rownames(mtcars))
)
```

```{r exercise6-solution}
rownames(mtcars)
```

<div class="mytask">
Create a new column called `brand` that contains the row names of `mtcars`. The `brand` column should be the last column in the data frame. Remember to output the data frame.
</div>

```{r exercise7, exercise = TRUE}
mtcars
```

```{r exercise7-check}
grade_this({
  mtcars$brand <- rownames(mtcars)
  pass_if_equal(mtcars)
})
```

```{r exercise7-solution}
mtcars$brand <- rownames(mtcars)
mtcars
```

## (E) Group summaries

<div class="mytask">
For every transmission level stored as `am` in `mtcars` (0 = automatic, 1 = manual), calculate the five number summary (min, Q1, Q2, Q3, max) of the horsepower, `hp`. Your output should be a data frame with columns `am`, `min`, `q1`, `med`, `q3` and `max`. Use `quantile` `type = 2` or equivalent for the calculation of Q1 and Q3. The `am` should be character. 
</div>

```{r exercise8, exercise = TRUE}
min <- ...
q1 <- ...(..., ..., function(x) quantile(x, 0.25, type = 2))
med <- ...
q3 <- ...
max <- ...
data.frame(am = names(min),
           min = min, q1 = q1, med = med, q3 = q3, max = max)
```

```{r exercise8-check}

grade_this({  
  min <- tapply(mtcars$hp, mtcars$am, min)
q1 <- tapply(mtcars$hp, mtcars$am, function(x) quantile(x, 0.25, type = 2))
med <- tapply(mtcars$hp, mtcars$am, function(x) quantile(x, 0.5, type = 2))
q3 <- tapply(mtcars$hp, mtcars$am, function(x) quantile(x, 0.75, type = 2))
max <- tapply(mtcars$hp, mtcars$am, max)
df <- data.frame(am = names(min),
                 min = min, q1 = q1, med = med, q3 = q3, max = max)
  pass_if_equal(df)
})
```

```{r exercise8-solution}
min <- tapply(mtcars$hp, mtcars$am, min)
q1 <- tapply(mtcars$hp, mtcars$am, function(x) quantile(x, 0.25, type = 2))
med <- tapply(mtcars$hp, mtcars$am, function(x) quantile(x, 0.5, type = 2))
q3 <- tapply(mtcars$hp, mtcars$am, function(x) quantile(x, 0.75, type = 2))
max <- tapply(mtcars$hp, mtcars$am, max)
data.frame(am = names(min),
           min = min, 
           q1 = q1, 
           med = med, 
           q3 = q3, 
           max = max)
```

## (F) Conditional values 

<div class="mytask">
For every cell in `mtcars`, set the value to 0 if its value is less than the average of its column.
</div>

```{r exercise9, exercise = TRUE}
for(nm in names(mtcars)) {
  mtcars[[nm]][... < mean(...)] <- 0
}
mtcars
```

```{r exercise9-check}
grade_this({
  for(nm in names(mtcars)) mtcars[[nm]][mtcars[[nm]] < mean(mtcars[[nm]])] <- 0
  pass_if_equal(mtcars)
})
```

```{r exercise9-solution}
for(nm in names(mtcars))  {
  mtcars[[nm]][mtcars[[nm]] < mean(mtcars[[nm]])] <- 0
}
mtcars
```

## (G) Rename column names

<div class="mytask">
Rename the column `wt` to `weight` and `hp` to `horsepower` for `mtcars`.
</div>

```{r exercise10, exercise = TRUE}
mtcars
```

```{r exercise10-check}
grade_this(
  pass_if_equal(dplyr::rename(mtcars, weight = wt, horsepower = hp))
)
```

```{r exercise10-solution}
nm <- names(mtcars)
names(mtcars)[nm=="hp"] <- "horsepower"
names(mtcars)[nm=="wt"] <- "weight"
mtcars
```

## (H) Rearrange columns

<div class="mytask">
Relocate `mpg` so that it appears after `vs` in `mtcars`.
</div>

```{r exercise11, exercise = TRUE}
mtcars
```

```{r exercise11-check}
grade_this(
  pass_if_equal(dplyr::relocate(mtcars, mpg, .after = vs))
)
```

```{r exercise11-solution}
nm <- names(mtcars)
mtcars[, nm[c(2:8, 1, 9:11)]]
```


