---
title: "Day 1 Exercise 1"
subtitle: "Knowing your `base` functions"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Revision of Base R functions.
---

```{r setup, include=FALSE, error = TRUE}
library(learnr)
if("gradethis" %in% installed.packages()) {
  gradethis::gradethis_setup()
} 
tutorial_options(exercise.timelimit = 60,
                 exercise.eval = FALSE)
knitr::opts_chunk$set(error = TRUE)
```

## Welcome

* This exercise is a revision of your knowledge in Base R. 
* You should try achieve the aim _without_ loading any extra packages.
* Note that answers can be achieved in multiple ways. 
* You'll working through the `mtcars` dataset.

```{r exercise0, exercise = TRUE}
mtcars
```


## (A) Subset with logical vectors

Extract every odd row from `mtcars`.

```{r exercise1, exercise = TRUE}

```

```{r exercise1-check}
grade_result(
  pass_if(~identical(.result, mtcars[c(TRUE, FALSE), ]))
)
```

```{r exercise2-solution}
mtcars[c(TRUE, FALSE), ]
```


<div id="exercise1-hint">
You can pass a logical vector to a `data.frame` object to subset rows or columns. What happens if the vector is shorter than the number of rows (or columns)?
</div>

## (B) Deleting a column or row

Delete the `wt` column from `mtcars`.

```{r exercise2, exercise = TRUE}

```

```{r exercise2-check}
grade_result(
  pass_if(~identical(.result, mtcars[c(TRUE, FALSE), ]))
)
```