---
title: "Day 2 Exercise 3"
subtitle: "Dealing with dates with `lubridate`"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Dealing with dates with lubridate
---

```{css, echo = FALSE}
.mytask {
  background: #82CAFA;
  border: 1.2px dashed #151B54;
  color: #151B54;
  padding: 10px;
  margin: auto;
  text-align: center;
}
.mytask:before {
  content: "Task";
  font-weight: bold;
}
```

```{r setup, include=FALSE, error = TRUE}
library(learnr)
library(tidyverse)
library(dwexercise)
library(lubridate)
if("gradethis" %in% installed.packages()) {
  gradethis::gradethis_setup()
} 
tutorial_options(exercise.timelimit = 60,
                 exercise.eval = FALSE)
knitr::opts_chunk$set(error = TRUE)
```

## Welcome

* You will be working with dates and times in these exercises.
* Note that answers can be achieved in multiple ways. 
* The `tidyverse` and `lubridate` packages and the datasets have been loaded for all exercises.

* The `oz_climate` data contains result from a survey about attitude towards climate change in Australia
* The `oz_climate_qbook` data contains the translation of the column label in `oz_climate` to the actual question asked

```{r exercise0A, exercise = TRUE}
data(oz_climate, package = "dwexercise")
oz_climate[1:3, ]
# ===only issue for online version===
# Note printing the whole data causes: 
# Error : pandoc document conversion failed with error 137
# looks like some issue with learnr?
```

```{r exercise0B, exercise = TRUE}
data(oz_climate_qbook, package = "dwexercise")
oz_climate_qbook
```


## (A) Computing time difference

<div class="mytask">
Compute the five number summary for the time taken to complete the survey for `oz_climate` by filling in `...` below.
</div>

```{r exercise1, exercise = TRUE}
oz_climate %>% 
  select(StartDate, EndDate) %>% 
  ... %>% 
  pull(Time) %>% 
  fivenum()
```

```{r exercise1-check}
grade_this({
  df <- oz_climate %>% 
    select(StartDate, EndDate) %>% 
    mutate(StartDate = mdy_hms(StartDate),
           EndDate = mdy_hms(EndDate),
           Time = EndDate - StartDate) %>% 
    pull(Time) %>% 
    fivenum()
  pass_if_equal(df)
})
```

```{r exercise1-solution}
oz_climate %>% 
  select(StartDate, EndDate) %>% 
  mutate(StartDate = mdy_hms(StartDate),
         EndDate = mdy_hms(EndDate),
         Time = EndDate - StartDate) %>% 
  pull(Time) %>% 
  fivenum()
```


## (B) Filter by date

<div class="mytask">
Filter `oz_climate` to surveys that were completed on or after August 13th 2011.
</div>

```{r exercise2, exercise = TRUE}
oz_climate %>% 
  mutate(...) %>% 
  filter(...) %>% 
  select(StartDate, EndDate)
```

```{r exercise2-check}
grade_this({
  df <- oz_climate %>% 
    mutate(EndDate = mdy_hms(EndDate)) %>% 
    filter(EndDate > ymd("20110813")) %>% 
    select(StartDate, EndDate)
  pass_if_equal(df$EndDate, x = .result$EndDate)
})
```

```{r exercise2-solution}
oz_climate %>% 
  mutate(EndDate = mdy_hms(EndDate)) %>% 
  filter(EndDate > ymd("20110813")) %>% 
  select(StartDate, EndDate)
```

## (C) Convert string to date, datetime or period

<div class="mytask">
Convert each string below to appropriate date, time or datetime objects.
</div>

```{r exercise3, exercise = TRUE}
...("May 2nd, 2014")
```

```{r exercise3-check}
grade_this({
  pass_if_equal(mdy("May 2nd, 2014"))
})
```

```{r exercise3-solution}
mdy("May 2nd, 2014")
```


```{r exercise4, exercise = TRUE}
...("Jan 30th (2019)")
```

```{r exercise4-check}
grade_this({
  pass_if_equal(mdy("Jan 30th (2019)"))
})
```

```{r exercise4-solution}
mdy("Jan 30th (2019)")
```

```{r exercise5, exercise = TRUE}
...("2020-12-29 08:30:27.243")
```

```{r exercise5-check}
grade_this({
  pass_if_equal(ymd_hms("2020-12-29 08:30:27.243"))
})
```

```{r exercise5-solution}
ymd_hms("2020-12-29 08:30:27.243")
```

```{r exercise6, exercise = TRUE}
...("Jun 28 2018 8:40AM")
```

```{r exercise6-check}
grade_this({
  pass_if_equal(mdy_hm("Jun 28 2018 8:40AM"))
})
```

```{r exercise6-solution}
mdy_hm("Jun 28 2018 8:40AM")
```

```{r exercise7, exercise = TRUE}
...("8:40") # 8 hours and 40 minutes
```

```{r exercise7-check}
grade_this({
  pass_if_equal(hm("8:40"))
})
```

```{r exercise7-solution}
hm("8:40")
```

## (D) When was it?

<div class="mytask">
What day of the week was Christmas in 1990? (Return the day abbreviated)
</div>

```{r exercise8, exercise = TRUE}

```

```{r exercise8-check}
grade_this({
  pass_if_equal(wday(dmy("25/12/90"), label = TRUE))
})
```

```{r exercise8-solution}
wday(dmy("25/12/90"), label = TRUE)
```

<div class="mytask">
What day of the year is it today?
</div>

```{r exercise9, exercise = TRUE}

```

```{r exercise9-check}
grade_this({
  pass_if_equal(yday(today()))
})
```

```{r exercise9-solution}
yday(today())
```

<div class="mytask">
What hour is it now in Singapore? Return as integer in 24-hour time.
</div>

```{r exercise10, exercise = TRUE}

```

```{r exercise10-check}
grade_this({
  pass_if_equal(hour(now(tzone = "Singapore")))
})
```

```{r exercise10-solution}
hour(now(tzone = "Singapore"))
```