---
title: "Day 2 Exercise 4"
subtitle: "Advanced data wrangling starring `purrr`"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Learning purrr and dplyr
---

```{css, echo = FALSE}
.mytask {
  background: #82CAFA;
  border: 1.2px dashed #151B54;
  color: #151B54;
  padding: 10px;
  margin: auto;
  text-align: center;
}
.mytask:before {
  content: "Task";
  font-weight: bold;
}
```

```{r setup, include=FALSE, error = TRUE}
library(learnr)
library(tidyverse)
library(dwexercise)
if("gradethis" %in% installed.packages()) {
  gradethis::gradethis_setup()
} 
tutorial_options(exercise.timelimit = 60,
                 exercise.eval = FALSE)
knitr::opts_chunk$set(error = TRUE)
```

## Welcome

* In this exercise you will practice using `purrr` as well as other data wrangling techniques you have learnt in the course.
* There are multiple ways to achieve the answers.
* Note that you don't necessary have to `purrr` for each question!
* The `tidyverse` package (which includes `purrr`) and the datasets have been loaded for all exercises.

```{r exercise0A, exercise = TRUE}
frog_signal
```

```{r exercise0B, exercise = TRUE}
oz_climate[1:3, ]
# ===only issue for online version===
# Note printing the whole data causes: 
# Error : pandoc document conversion failed with error 137
# looks like some issue with learnr?
```

## (A) Data set characteristics 

<div class="mytask">
What are class type for each column in `frog_signal` data? 
</div>


```{r exercise1, exercise = TRUE}

```

```{r exercise1-check}
grade_this({
  pass_if_equal(map_chr(frog_signal, class))
})
```

```{r exercise1-solution}
frog_signal %>% 
  map_chr(class)
```


<div class="mytask">
What are the number of distinct values in each column in `oz_climate`? (Return number as integer)
</div>


```{r exercise2, exercise = TRUE}

```

```{r exercise2-check}
grade_this({
  pass_if_equal(map_int(oz_climate, n_distinct))
})
```

```{r exercise2-solution}
map_int(oz_climate, n_distinct)
```



## (B) Principle Component Analysis

<div class="mytask">
Subset to only numerical variables in `frog_signal`. Remove any rows with missing values. And then normalise every numerical variable such that its sample mean is 0 and sample standard deviation is 1. And then run a principle component analysis.
</div>


```{r exercise3, exercise = TRUE}
frog_signal %>% 
  ... %>% 
  drop_na() %>% 
  scale() %>% 
  prcomp()
```

```{r exercise3-check}
grade_this({
  df <- frog_signal %>% 
    select(where(is.numeric)) %>% 
    drop_na() %>% 
    scale() %>% 
    prcomp()
  pass_if_equal(df)
})
```

```{r exercise3-solution}
frog_signal %>% 
  select(where(is.numeric)) %>% 
  drop_na() %>% 
  scale() %>% 
  prcomp()
```

## (C) Simulation study: misspecified model

* Let's simulate a data with 200 observations from a simple linear model with intercept is 1 and slope is 2 and where the error distribution is from a $t_3$ distribution.  


```{r exercise4-show, exercise = TRUE}
set.seed(1)
sim <- tibble(1:200) %>% 
  mutate(x = runif(n(), 0, 1), # some independent covariates
         y = 1 + 2 * x + rt(n(), 3)) # intercept = 1, slope = 2

ests <- coef(lm(y ~ x, data = sim))
tibble(intercept = ests[1], slope = ests[2]) # how close are these to the true values?
```

<div class="mytask">
Now repeat the above simulation 100 times. Compute the mean and sample standard deviation of the absolute difference between estimated intercept and slope to its true value.
</div>

```{r exercise4, exercise = TRUE}
set.seed(1)
nsim <- 100
...
```

```{r exercise4-solution}
set.seed(1)
nsim <- 100
df <- map_dfr(1:nsim, ~{
  sim <- tibble(1:200) %>% 
    mutate(x = runif(n(), 0, 1), # some independent covariates
           y = 1 + 2 * x + rt(n(), 3)) 
  ests <- coef(lm(y ~ x, data = sim))
  tibble(intercept = ests[1], slope = ests[2], sim =.x)
})

df %>% 
  mutate(intercept = abs(intercept - 1),
         slope = abs(slope - 2)) %>% 
  select(-sim) %>% 
  summarise(across(intercept:slope, 
                   list(mean = mean,
                        sd = sd),
                   .names = "{.col}_{.fn}")) 
```

* The mean and sample standard deviation of the absolute difference between estimated intercept and slope to its true value is shown below.
* How closely does it match up with yours?

```{r exercise4-print, echo = FALSE, message = FALSE, warning = FALSE}
set.seed(1)
df <- map_dfr(1:100, ~{
  sim <- tibble(1:200) %>% 
    mutate(x = runif(n(), 0, 1), # some independent covariates
           y = 1 + 2 * x + rt(n(), 3)) 
  ests <- coef(lm(y ~ x, data = sim))
  tibble(intercept = ests[1], slope = ests[2], sim =.x)
})

df %>% 
  mutate(intercept = abs(intercept - 1),
         slope = abs(slope - 2)) %>% 
  select(-sim) %>% 
  summarise(across(intercept:slope, 
                   list(mean = mean,
                        sd = sd),
                   .names = "{.col}_{.fn}")) 
```