---
title: "Day 2 Exercise 1"
subtitle: "Relational data wrangling starring `janitor` and `broom`"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Relational data wrangling starring janitor and broom
---

```{css, echo = FALSE}
.mytask {
  background: #82CAFA;
  border: 1.2px dashed #151B54;
  color: #151B54;
  padding: 10px;
  margin: auto;
  text-align: center;
}
.mytask:before {
  content: "Task";
  font-weight: bold;
}
```

```{r setup, include=FALSE, error = TRUE}
library(learnr)
library(tidyverse)
library(janitor)
library(broom)
library(dwexercise)
library(agridat)
if("gradethis" %in% installed.packages()) {
  gradethis::gradethis_setup()
} 
tutorial_options(exercise.timelimit = 60,
                 exercise.eval = FALSE)
knitr::opts_chunk$set(error = TRUE)
```

## Welcome

* These exercises have get you to work with combining datasets and appending of other information (e.g. totals, percentages, or model output) using `janitor` and `broom`.
* Note that answers can be achieved in multiple ways. 
* The `tidyverse`, `janitor`, and `broom` packages and the datasets have been loaded for all exercises.

* The data `census_2016_G04` contains the number of people for each age and sex combination by Australian states and territories from the 2016 census.

```{r exercise0A, exercise = TRUE}
data(census_2016_G04, package = "dwexercise")
census_2016_G04
```

* The data `census_2016_STE` contains the dictionary for STE code to the names of the corresponding states and territories. 

```{r exercise0B, exercise = TRUE}
data(census_2016_STE, package = "dwexercise")
census_2016_STE
```

* The data `nass.wheat` and `nass.soybean` contains the annual wheat and soybean yield (in bushels/acre) for each state.

```{r exercise0C, exercise = TRUE}
data(nass.wheat, package = "agridat")
nass.wheat
```


```{r exercise0D, exercise = TRUE}
data(nass.soybean, package = "agridat")
nass.soybean
```


## (A) Pivot data once again

<div class="mytask">
Transform the data `census_2016_G04` such that it is in a long format as shown below. Note that the transformed data ignores the aggregated versions of other columns and people who are 80 or over 80. Save this output as `dfA`. If you have a hard time figuring out the string pattern, see the hint.
</div>

```{r exercise1, exercise = TRUE}
dfA <- census_2016_G04 %>% 
  ...(STE_CODE_2016 | matches(...)) %>% 
  ...(...,
      ... = c("age", "sex"),
      ... = ...,
      names_transform = list(age = as.integer),
      ... = "count")
```

```{r exercise1-check}
grade_this({
  df <- census_2016_G04 %>% 
    select(STE_CODE_2016 | matches("^Age_yr_[0-9]{1,2}_[MF]$")) %>% 
    pivot_longer(-STE_CODE_2016,
                 names_to = c("age", "sex"),
                 names_pattern = "^Age_yr_([0-9]{1,2})_([MF])",
                 names_transform = list(age = as.integer),
                 values_to = "count")
  pass_if_equal(df)
})
```

```{r exercise1-solution}
dfA <- census_2016_G04 %>% 
  select(STE_CODE_2016 | matches("^Age_yr_[0-9]{1,2}_[MF]$")) %>% 
  pivot_longer(-STE_CODE_2016,
               names_to = c("age", "sex"),
               names_pattern = "^Age_yr_([0-9]{1,2})_([MF])",
               names_transform = list(age = as.integer),
               values_to = "count")
```

* Your output data should be like below

```{r exercise1-print, echo = FALSE}
census_2016_G04 %>% 
  select(STE_CODE_2016 | matches("^Age_yr_[0-9]{1,2}_[MF]$")) %>% 
  pivot_longer(-STE_CODE_2016,
               names_to = c("age", "sex"),
               names_pattern = "^Age_yr_([0-9]{1,2})_([MF])",
               names_transform = list(age = as.integer),
               values_to = "count")
```

<details><summary>Hint</summary>
The `names_pattern` regular expression should be `"^Age_yr_([0-9]{1,2})_([MF])"`.
</details>

## (B) Replace with dictionary values

```{r exercise2-setup, echo = FALSE}
dfA <- census_2016_G04 %>% 
  select(STE_CODE_2016 | matches("^Age_yr_[0-9]{1,2}_[MF]$")) %>% 
  pivot_longer(-STE_CODE_2016,
               names_to = c("age", "sex"),
               names_pattern = "^Age_yr_([0-9]{1,2})_([MF])",
               names_transform = list(age = as.integer),
               values_to = "count")
```

<div class="mytask">
Given the transformed data from (A), now match up with the data dictionary `census_2016_STE` to get the actual names of the states and territories; and then transform the data to a wide format as shown below. 
</div>

```{r exercise2, exercise = TRUE}
dfB <- dfA %>% 
  ...(census_2016_STE, by = ...) %>% 
  select(-STE_CODE_2016, -name) %>% 
  ...
```

```{r exercise2-check}
grade_this({
  dfA <- census_2016_G04 %>% 
  select(STE_CODE_2016 | matches("^Age_yr_[0-9]{1,2}_[MF]$")) %>% 
  pivot_longer(-STE_CODE_2016,
               names_to = c("age", "sex"),
               names_pattern = "^Age_yr_([0-9]{1,2})_([MF])",
               names_transform = list(age = as.integer),
               values_to = "count")

  df <- dfA %>% 
    left_join(census_2016_STE, by = c("STE_CODE_2016" = "code")) %>% 
    select(-STE_CODE_2016, -name) %>% 
    pivot_wider(names_from = abbrv,
                values_from = count)
  pass_if_equal(df)
})
```

```{r exercise2-solution}
dfB <- dfA %>% 
  left_join(census_2016_STE, by = c("STE_CODE_2016" = "code")) %>% 
  select(-STE_CODE_2016, -name) %>% 
  pivot_wider(names_from = abbrv,
              values_from = count)
```

* Your output data should be like below

```{r exercise2-print, echo = FALSE}
dfA <- census_2016_G04 %>% 
  select(STE_CODE_2016 | matches("^Age_yr_[0-9]{1,2}_[MF]$")) %>% 
  pivot_longer(-STE_CODE_2016,
               names_to = c("age", "sex"),
               names_pattern = "^Age_yr_([0-9]{1,2})_([MF])",
               names_transform = list(age = as.integer),
               values_to = "count")

dfA %>% 
  left_join(census_2016_STE, by = c("STE_CODE_2016" = "code")) %>% 
  select(-STE_CODE_2016, -name) %>% 
  pivot_wider(names_from = abbrv,
              values_from = count)

```

## (C) Adorn with totals

```{r exercise3-setup, echo = FALSE}
dfA <- census_2016_G04 %>% 
  select(STE_CODE_2016 | matches("^Age_yr_[0-9]{1,2}_[MF]$")) %>% 
  pivot_longer(-STE_CODE_2016,
               names_to = c("age", "sex"),
               names_pattern = "^Age_yr_([0-9]{1,2})_([MF])",
               names_transform = list(age = as.integer),
               values_to = "count")

dfB <- dfA %>% 
  left_join(census_2016_STE, by = c("STE_CODE_2016" = "code")) %>% 
  select(-STE_CODE_2016, -name) %>% 
  pivot_wider(names_from = abbrv,
              values_from = count)
```

<div class="mytask">
Given the transformed data from (B), add a column and row that show the total population by row (sex and age) and column (state). For each cell that has the count, append a percentage by each sex and age group as shown below.
</div>

```{r exercise3, exercise = TRUE}
dfB %>% 
  adorn...(c("row", "col")) %>% 
  ...("row") %>% 
  ...(digits = 2) %>%
  ...(position = "front")
```

```{r exercise3-check}
grade_this({
  dfA <- census_2016_G04 %>% 
  select(STE_CODE_2016 | matches("^Age_yr_[0-9]{1,2}_[MF]$")) %>% 
  pivot_longer(-STE_CODE_2016,
               names_to = c("age", "sex"),
               names_pattern = "^Age_yr_([0-9]{1,2})_([MF])",
               names_transform = list(age = as.integer),
               values_to = "count")

dfB <- dfA %>% 
  left_join(census_2016_STE, by = c("STE_CODE_2016" = "code")) %>% 
  select(-STE_CODE_2016, -name) %>% 
  pivot_wider(names_from = abbrv,
              values_from = count)

  df <- dfB %>% 
    adorn_totals(c("row", "col")) %>% 
    adorn_percentages("row") %>% 
    adorn_pct_formatting(digits = 2) %>%
    adorn_ns(position = "front")
  pass_if_equal(df)
})
```

```{r exercise3-solution}
dfB %>% 
  adorn_totals(c("row", "col")) %>% 
  adorn_percentages("row") %>% 
  adorn_pct_formatting(digits = 2) %>%
  adorn_ns(position = "front")
```

* Your output data should be like below

```{r exercise3-print, echo = FALSE}
dfA <- census_2016_G04 %>% 
  select(STE_CODE_2016 | matches("^Age_yr_[0-9]{1,2}_[MF]$")) %>% 
  pivot_longer(-STE_CODE_2016,
               names_to = c("age", "sex"),
               names_pattern = "^Age_yr_([0-9]{1,2})_([MF])",
               names_transform = list(age = as.integer),
               values_to = "count")
dfA %>% 
  left_join(census_2016_STE, by = c("STE_CODE_2016" = "code")) %>% 
  select(-STE_CODE_2016, -name) %>% 
  pivot_wider(names_from = abbrv,
              values_from = count)

dfB %>% 
  adorn_totals(c("row", "col")) %>% 
  adorn_percentages("row") %>% 
  adorn_pct_formatting(digits = 2) %>%
  adorn_ns(position = "front")
```



## (D) Join, nest and model


```{r exercise4-plot, echo = FALSE, message = FALSE}
nass.wheat %>% 
  inner_join(nass.soybean, by = c("year", "state"),
            suffix = c("_wheat", "_soybean")) %>% 
  ggplot(aes(yield_wheat, yield_soybean)) + 
  facet_wrap(~state) + geom_point() + 
  geom_smooth(method = "lm") + 
  labs(x = "Wheat", y = "Soybean", title = "Yield")
```

<div class="mytask">
Above is the plot of the soy bean yield against wheat yield for each state from `nass.soybean` and `nass.wheat` data. The blue line that overlays the plot is a simple linear model. Construct a data frame that contains the intercept and slope and its standard errors of each model with rows arranged alphabetically by state. 
</div>

```{r exercise4, exercise = TRUE}
nass.wheat %>% 
  ...(nass.soybean, by = c("year", "state"),
            suffix = c("_wheat", "_soybean")) %>% 
  ...(state) %>% 
  nest() %>% 
  # fit the model and get the estimates and its std.err
  mutate(model = list(lm(yield_wheat ~ yield_soybean, data = ...)),
         ests = list(tidy(...))) %>% 
  unnest(ests) %>% 
  # remove unnecessary statistics
  select(-c(statistic, p.value)) %>% 
  # convert to the final form as required
  mutate(term = ifelse(term=="yield_soybean", "slope", term)) %>% 
  ...(names_from = ...,
      values_from = ...) %>% 
  clean_names()
```

```{r exercise4-check}
grade_this({
  df <- nass.wheat %>% 
    inner_join(nass.soybean, by = c("year", "state"),
              suffix = c("_wheat", "_soybean")) %>% 
    group_by(state) %>% 
    nest() %>% 
    # fit the model and get the estimates and its std.err
    mutate(model = list(lm(yield_wheat ~ yield_soybean, data = data[[1]])),
           ests = list(tidy(model[[1]]))) %>% 
    unnest(ests) %>% 
    # remove unnecessary statistics
    select(-c(statistic, p.value)) %>% 
    # convert to the final form as required
    mutate(term = ifelse(term=="yield_soybean", "slope", term)) %>% 
    pivot_wider(names_from = term,
                values_from = estimate:std.error) %>% 
    clean_names()
  pass_if_equal(df$estimate_slope, x = .result[[names(.result)[str_detect(names(.result), "slope") & !str_detect(names(.result), "error")]]])
})
```

```{r exercise4-solution}
nass.wheat %>% 
  inner_join(nass.soybean, by = c("year", "state"),
            suffix = c("_wheat", "_soybean")) %>% 
  group_by(state) %>% 
  nest() %>% 
  # fit the model and get the estimates and its std.err
  mutate(model = list(lm(yield_wheat ~ yield_soybean, data = data[[1]])),
         ests = list(tidy(model[[1]]))) %>% 
  unnest(ests) %>% 
  # remove unnecessary statistics
  select(-c(statistic, p.value)) %>% 
  # convert to the final form as required
  mutate(term = ifelse(term=="yield_soybean", "slope", term)) %>% 
  pivot_wider(names_from = term,
              values_from = estimate:std.error) %>% 
  clean_names()
```

* Your output data should be like below

```{r exercise4-print, echo = FALSE}
nass.wheat %>% 
  inner_join(nass.soybean, by = c("year", "state"),
            suffix = c("_wheat", "_soybean")) %>% 
  group_by(state) %>% 
  nest() %>% 
  # fit the model and get the estimates and its std.err
  mutate(model = list(lm(yield_wheat ~ yield_soybean, data = data[[1]])),
         ests = list(tidy(model[[1]]))) %>% 
  unnest(ests) %>% 
  # remove unnecessary statistics
  select(-c(statistic, p.value)) %>% 
  # convert to the final form as required
  mutate(term = ifelse(term=="yield_soybean", "slope", term)) %>% 
  pivot_wider(names_from = term,
              values_from = estimate:std.error) %>% 
  clean_names()
```